---
title: "Time Series Analysis and Forecasting of China's Raw Coal Production"
author: "Qu Youyang"
date: "2024-06"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
---

```{r setup, include = FALSE}
# Global options for the report
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

1. Project Overview
This project focuses on modeling and forecasting China's monthly raw coal production (Aug 2000 - Aug 2024). We employ seasonal ARIMA models and Holt-Winters exponential smoothing, comparing automated selection with manual stepwise modeling.

2. Environment Setup
```{r}
# Load necessary libraries
library(tsdiy)
library(forecast)
library(readxl)
library(zoo)
library(ggplot2)

# Verify tsdiy version
if (packageVersion("tsdiy") < "1.6") {
  stop("Please update tsdiy to version 1.6")
}
```

3. Data Loading and Preprocessing
The dataset is sourced from the National Bureau of Statistics. We handle missing values using spline interpolation and apply log-transformation to stabilize variance.
```{r}
# Load data (ensure 'data.xlsx' is in the working directory)
# Change the path below if your filename is different
data_raw <- read_xlsx("data.xlsx") 

coal <- ts(data_raw$`原煤产量当期值(万吨)`, start = c(2000, 8), end = c(2024, 8), frequency = 12)

# EDA: Visualizing Trend and Seasonality
par(mfrow = c(1, 3), mar = c(2.5, 2.5, 2, 0.5), mgp = c(1.5, 0.5, 0))
ts.plot(coal, xlab = "Year", ylab = "10k Tons", main = "Original Series")
ans <- MAfilter(coal, k = 12)
lines(c(time(coal)), ans$trend, col = "red")

# Residual Analysis
plot(ans$trend, ans$remainder, xlab = "Trend", ylab = "Remainder", main = "Residuals")
sacf <- SACF(ans$remainder, lag.max = 50)

# Preprocessing: Log and Spline Interpolation
logcoalnew <- na.spline(log(coal))
tslogcoalnew <- ts(logcoalnew, start = c(2000, 8), freq = 12)
```

4. ARIMA Modeling
4.1 Automated ARIMA (forecast package)
```{r}
model_auto <- auto.arima(tslogcoalnew)
summary(model_auto)
```

4.2 Stepwise ARIMA (tsdiy package)
Comparison of manually identified seasonal ARIMA models.
```{r}
# Seasonal Differencing (D=1)
diff_coal <- diff(tslogcoalnew, lag = 12)

# Candidate 1: ARIMA(1,0,1)(0,1,1)[12]
fit_manual1 <- ARIMA(tslogcoalnew, pdq = c(1,0,1), PDQ = c(0,1,1), s=12, include.drift = FALSE)

# Candidate 2: ARIMA(0,1,1)(0,1,1)[12]
fit_manual2 <- ARIMA(tslogcoalnew, pdq = c(0,1,1), PDQ = c(0,1,1), s=12)
```

5. Holt-Winters Exponential Smoothing
```{r}
holt_fit <- HWS(tslogcoalnew, type = 3) # Additive Model
plot(holt_fit, main = "Holt-Winters Fitting")
```

6. Forecasting via Bootstrap
Implementing Residual Bootstrap for future 12-period production.
```{r}
# Using ARIMA(0,1,1)(0,1,1)[12] as an example
IMA <- fit_manual2
b <- IMA$coef[1,1]
bs <- IMA$coef[1,2]
p <- tsdiy:::polyprod(c(1,b), c(1, rep(0,11), bs))
b_vec <- p[2:14]

N <- 1000
x.forecast <- matrix(nrow = 12, ncol = N)

for (i in 1:N) {
  e.bootstrap <- sample(IMA$residuals, size = 12, replace = TRUE)
  z.forecast <- maq(e = e.bootstrap, b = b_vec, init = tail(IMA$residuals, length(b_vec)), mu = 0)
  # Inverse differencing
  y.forecast <- rwd(e = z.forecast, init = tail(diff(tslogcoalnew, 12), 1), theta = 0.00)
  x.forecast[, i] <- rwDs(e = y.forecast, D = 1, s = 12, init = tail(tslogcoalnew, 12))
}

# Final Forecasting Results
forecast_mean <- rowMeans(x.forecast)
forecast_lower <- apply(x.forecast, 1, quantile, probs = 0.025)
forecast_upper <- apply(x.forecast, 1, quantile, probs = 0.975)

# Compare with real 2024 Q4 data (Log scale)
real_log_2024 <- log(c(41445.9, 41180.3, 42798.4, 43884.8))
```

7. Summary
The results indicate that manual stepwise ARIMA modeling with Residual Bootstrap provides a more parsimonious and stable prediction for China's raw coal output compared to automated black-box methods.